\pagestyle{empty}
\vspace*{\fill}%

\section*{Abstract}

In this thesis, we develop a module and configuration system for the Jolie
programming language. The features of these systems allow for new ways in which
to create and distribute Jolie services. This makes it possible to change
freely between binding to external services to internally embedded services
entirely from configuration, and vice versa. Additionally, the configuration
system introduces a notion of interface parametricity that can be used to
develop generic Jolie services.

The module and configuration system lay the groundwork for the Jolie Package
Manager (JPM). JPM is described and its implementation details are discussed in
the second part of this thesis. JPM provides a helpful tool which makes it
easier to develop and share Jolie services with other developers.

\section*{Resume}

I dette speciale udvikler vi et modul- og konfigurationssystem til
programmeringssproget Jolie. Funktionerne som dette system tilbyder, giver nye
muligheder for at udvikle og dele Jolie servicer. Dette gør det muligt, at frit
skifte, frem og tilbage, mellem eksterne bindinger til servicer og binde til en
indlejret intern service. Desuden introducerer konfigurationsystemet en form
for ``interface parametricity'', som tilllader udviklingen af generiske Jolie
servicer.

Modul- og konfigurationssystemet er grundstenen, som tillader udviklingen af
JPM (Jolie Package Manager). JPM beskrives og interessante implementerings
detaljer diskuteres i den anden halvdel af dette speciale. JPM fremsætter en
række værktøjer, hvilket gør det nemmere at udvikle og dele Jolie servicer med
andre udviklere.

\vspace*{\fill}%
