\section{JPM Manifest Specification}\label{package-specification}

\hypertarget{name}{\subsubsection{name}\label{name}}

\textbf{Field name:} \texttt{name}

\textbf{Optional:} false

\textbf{Type:} \texttt{string}

\textbf{Description:} The \texttt{name} property uniquely defines a
package in a registry. Every registry must only contain a single package
with a given name.

\textbf{Rules:}

\begin{itemize}
\tightlist
\item
  The name of a package is \emph{not} case-sensitive
\item
  The length of a name is less than 255 characters
\item
  Names are US-ASCII
\item
  Names may only contain unreserved URI characters (see section 2.3 of
  \href{https://www.ietf.org/rfc/rfc3986.txt}{RFC 3986})
\end{itemize}

If any of these rules are broken the JPM tool should complain when
\emph{any} command is invoked. Similarly a registry should reject any
such package.

\hypertarget{version}{\subsubsection{version}\label{version}}

\textbf{Field name:} \texttt{version}

\textbf{Optional:} false

\textbf{Type:} \texttt{string}

\textbf{Description:} This property describes the current version of
this package.

\textbf{Rules:}

\begin{itemize}
\tightlist
\item
  The version string must be a valid SemVer 2.0.0 string (see
  http://semver.org/spec/v2.0.0.html)
\end{itemize}

\hypertarget{license}{\subsubsection{license}\label{license}}

\textbf{Field name:} \texttt{property\_name}

\textbf{Optional:} false

\textbf{Type:} \texttt{string}

\textbf{Description:} Describes the license that this package is under.

\textbf{Rules:}

\begin{itemize}
\tightlist
\item
  Must be a valid identifier. See https://spdx.org/licenses/
\end{itemize}

\hypertarget{authors}{\subsubsection{authors}\label{authors}}

\textbf{Field name:} \texttt{authors}

\textbf{Optional:} false

\textbf{Type:}
\texttt{string\textbar{}array\textless{}string\textgreater{}}

\textbf{Description:} Describes the authors of this package

\textbf{Rules:}

\begin{itemize}
\tightlist
\item
  The array must contain at least a single entry
\item
  Each entry should follow this grammar:
\end{itemize}

\begin{verbatim}
name ["<" email ">"] ["(" homepage ")"]
\end{verbatim}

\hypertarget{private}{\subsubsection{private}\label{private}}

\textbf{Field name:} \texttt{private}

\textbf{Optional:} true

\textbf{Type:} \texttt{boolean}

\textbf{Description:} Describes if this package should be considered
private. If a package is private it cannot be published to the
``public'' repository.

\textbf{Rules:}

\begin{itemize}
\tightlist
\item
  By default this property has the value of \texttt{true} to avoid
  accidential publishing of private packages.
\end{itemize}

\hypertarget{main}{\subsubsection{main}\label{main}}

\textbf{Field name:} \texttt{main}

\textbf{Optional:} true

\textbf{Type:} \texttt{string}

\textbf{Description:} Describes the main file of a package.

\textbf{Rules:}

\begin{itemize}
\tightlist
\item
  The value is considered to be a relative file path from the package
  root.
\end{itemize}

\hypertarget{dependencies}{\subsubsection{dependencies}\label{dependencies}}

\textbf{Field name:} \texttt{dependencies}

\textbf{Optional:} true

\textbf{Type:} \texttt{array\textless{}dependency\textgreater{}}

\textbf{Description:} Contains an array of dependencies. See the
``dependency'' sub-section for more details.

\textbf{Rules:}

\begin{itemize}
\tightlist
\item
  If the property is not listed, a default value of an empty array
  should be used
\end{itemize}

\hypertarget{dependency}{\subsubsection{dependency}\label{dependency}}

\textbf{Type:} \texttt{object}

\textbf{Description:} A dependency describes a single dependency of a
package. This points to a package at a specific point on a specific
registry.

\hypertarget{name-1}{\paragraph{dependency.name}\label{name-1}}

\textbf{Field name:} \texttt{name}

\textbf{Optional:} false

\textbf{Type:} \texttt{string}

\textbf{Description:} Describes the name of the dependency. This refers
to the package name, as defined earlier.

\textbf{Rules:} A dependency name follows the exact same rules as a
package name.

\hypertarget{version-1}{\paragraph{dependency.version}\label{version-1}}

\textbf{Field name:} \texttt{version}

\textbf{Optional:} false

\textbf{Type:} \texttt{string}

\textbf{Description:} Describes the version to use

\textbf{Rules:}

\begin{itemize}
\tightlist
\item
  Must be a valid SemVer 2.0.0 string
\item
  (This property follows the same rules as the package version does)
\end{itemize}

\hypertarget{registry}{\paragraph{dependency.registry}\label{registry}}

\textbf{Field name:} \texttt{registry}

\textbf{Optional:} true

\textbf{Type:} \texttt{string}

\textbf{Description:} This describes the exact registry to use. If no
registry is listed the ``public'' registry will be used.

\textbf{Rules:}

\begin{itemize}
\tightlist
\item
  The value of this property must be a valid registry as listed in the
  \texttt{registries} property.
\end{itemize}

\hypertarget{registries}{\subsubsection{registries}\label{registries}}

\textbf{Field name:} \texttt{registries}

\textbf{Optional:} true

\textbf{Type:} \texttt{array\textless{}registry\textgreater{}}

\textbf{Description:} Contains an array of known registries. See the
registry sub-section for more details.

\textbf{Rules:}

\begin{itemize}
\tightlist
\item
  This property contains an implicit entry which points to the public
  registry. This registry is named ``public''.
\end{itemize}

\hypertarget{registry-1}{\subsubsection{registry}\label{registry-1}}

\textbf{Type:} \texttt{object}

\textbf{Description:} A registry describes a single JPM registry. A JPM
registry is where the package manager can locate a package, and also
request a specific version of a package.

\hypertarget{name-2}{\paragraph{registry.name}\label{name-2}}

\textbf{Field name:} \texttt{name}

\textbf{Optional:} false

\textbf{Type:} \texttt{string}

\textbf{Description:} This property uniquely identifies the registry.

\textbf{Rules:}

\begin{itemize}
\tightlist
\item
  A name cannot be longer than 1024 characters
\item
  The name cannot be ``public''
\item
  No two registries may have the same name
\end{itemize}

\hypertarget{location}{\paragraph{registry.location}\label{location}}

\textbf{Field name:} \texttt{location}

\textbf{Optional:} false

\textbf{Type:} \texttt{string}

\textbf{Description:} Describes the location of the registry.

\textbf{Rules:}

\begin{itemize}
\tightlist
\item
  Must be a valid Jolie location string (e.g.
  ``socket://localhost:8080'')
\end{itemize}
