\section{Lifetime Hooks}
\label{sec:lifetime_hooks}

A lifetime hook is a script that ``hooks'' onto certain lifetime events that
occur during normal use of the package manager. These scripts allows the
package developer to augment the package manager, and potentially affect the
work it performs.

This is a feature which is available in many other types of software, for
example the popular version control system Git provides such a
feature\autocite{chacon2009pro}.

All hooks are performed on the client-side, and never on the server side (i.e.
on registries). Only the hooks of the current package will be run, the
hooks of a dependency will not be executed. Hooks that run before a certain
action, denoted by a \verb!pre-! prefix, may stop the corresponding command
from being executed.  These scripts may do so by returning a non-zero exit
code.

These features become useful for augmenting the workflow with in-house
conventions. This may for example include ensuring that, for example, tests
pass before publishing a new version. Using a simple process interface allows
for the user to write these scripts in whichever technology they choose.

The hooks are defined in the ``events'' section of the package manifest.
Listing \ref{lst:hook} shows an example hook, which runs before publishing
which ensure that all tests pass.

\begin{listing}[H]
\begin{minted}{javascript}
{
    // ...
    "events": {
        "pre-publish": "./run_tests.sh"
    }
}
\end{minted}

\caption{Defining a lifetime hook, which runs the script
    \mintinline{text}{run_tests.sh} before publishing the package to the
        registry}

\label{lst:hook}

\end{listing}

The following are the hooks that JPM supports:

\begin{itemize}

\item \verb!pre-start!: Script is run right before the \verb!start! command is
    invoked on a service.

\item \verb!post-start!: Run right before the \verb!start! command
    terminates. There is no guarantee that this script is run (i.e. if the
    service was forcefully terminated).

\item \verb!pre-install!: Runs before the \verb!install! command is invoked.

\item \verb!post-install!: Runs after the \verb!install! command has finished.

\item \verb!pre-publish!: Runs before the \verb!publish! command is invoked.

\item \verb!post-publish!: Runs after the \verb!publish! command has finished.

\end{itemize}
